---
title: "Parameter Estimation – Wild Fish Catch"
author: "Bri Baker"
date: "3/2/2021"
output:
  html_document:
    theme: cosmo
    code_folding: hide
    highlight: kate
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(nlstools)

```
## Summary


```{r data import}
fish <- read_csv(here("data/fish_catch.csv"),
                 skip = 2,
                 col_names = TRUE) %>% 
  clean_names() %>% 
  na.omit() %>% 
  rowid_to_column("years") %>% 
  mutate(years = years-1) %>% 
  mutate_if(is.character, as.numeric,) %>% 
  select(years, wild_catch)
```


## Data Exploration

Exploratory plot of wild caught fish.

```{r explore}
ggplot(fish, aes(years, wild_catch))+
  geom_point()+
  labs(x = "years", y = "wild caught")+
  scale_y_continuous(breaks = seq(0, 100, 10))+
  theme_minimal()

```

Logistic growth curve which can be modeled as $P(t)=\frac{K}{1+Ae^{-kt}}$($K$ = carrying capacity, $A$ = $\frac{K-P_0}{P_0}$, $k$ = growth rate constant).

## Initial parameter estimation

```{r parameters}
K <-  90 # carrying capacity at ~90
A <- (K-10)/10 # Po ~10

exp_fish <- fish %>% 
  filter(years < 38) %>% # filter for exp
  mutate(ln_pop = log(wild_catch)) # log transform

lm_fish <- lm(ln_pop ~ years, data = exp_fish)

r <- lm_fish$coefficients[[2]]

```

Estimated starting parameters: $K$ = `r K`, $A$ = `r A`, $k$ = `r r`.


## Nonlinear least squares parameter estimation

```{r}
fish_nls <- nls(wild_catch ~ K/(1 + A*exp(-r*years)),
                data = fish,
                start = list(K = 90, A = 8, r = 0.035),
                trace = TRUE)
```






## Citations

Earth Policy Institute (2010). Global wild fish catch and aquaculture production, compiled by Earth Policy Institute with 1950-2010 from U.N. Food and Agriculture Organization (FAO), Global Capture Production and Global Aquaculture Production, electronic databases, at www.fao.org/fishery/topic/16140/en.

Firke, Sam (2020). janitor: Simple Tools for Examining and Cleaning Dirty Data. R package version 2.0.1. https://CRAN.R-project.org/package=janitor

Müller, Kirill (2020). here: A Simpler Way to Find Your Files. R package version 1.0.0. https://CRAN.R-project.org/package=here

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686. https://doi.org/10.21105/joss.01686
