---
title: 'Bootstrapped Confidence Interval for Proportions'
author: "Bri Baker"
date: "3/3/2021"
output:
  html_document:
    theme: cosmo
    code_folding: hide
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

library(tidyverse)
library(here)
library(boot)
```

# Summary

In a 2014 survey of UCSB community members, 22 out of 36 surveyed UCSB community members (61%) identifying as nonbinary or genderqueer responded that they had personally experienced “exclusionary, offensive, hostile or intimidating conduct” (Rankin & Associates Consulting, 2014). 

Here, I utilize bootstrapping to determine a 95% confidence interval for this statistic, based off of the observed data (n = 36).


```{r data}

hurt <- rep(c(1 , 0), times = c(22, 14)) # 1 for yes, 0 for no


```

```{r bootstrap}

ratio_hurt <- function(x,i){
  sum(x[i])/36 # calculate ratio hurt
}

hurtboot_10k <- boot(
  hurt, # use original data
  statistic = ratio_hurt, # find ratio hurt
  R = 10000 # bootstrap 10k times
)

hurt_ci <- boot.ci(hurtboot_10k, conf = 0.95)

ci_vec <- hurt_ci$normal

```

```{r histogram}

boot_ratio <- as_tibble(hurtboot_10k$t)

ggplot(boot_ratio, aes(V1)) +
  geom_histogram(bins = 20) +
  labs(title = "Proportion of nonbinary and genderqueer people hurt by conduct at UCSB",
       subtitle = "Bootstrapping analysis of 2014 survey data",
       x = "Proportion",
       y = "Count",
       caption = "Bri Baker, 2021\n(Rankin & Associates Consulting, 2014)") +
  theme_minimal()

```

The confidence interval when these data are bootstrapped 10,000 times is `r round(ci_vec[2], 2)`-`r round(ci_vec[3], 3)` indicating that there is a 95% chance that 95 of 100 calculated confidence intervals will contain the actual population parameter in that range. 



## Citations

Canty, Angelo and Brian Ripley (2020). boot: Bootstrap R (S-Plus) Functions. R package version 1.3-25.

Müller, Kirill (2020). here: A Simpler Way to Find Your Files. R package version 1.0.0. https://CRAN.R-project.org/package=here

Rankin & Associates Consulting (2014). UCSB Campus Climate Project Final Report. http://campusclimate.ucop.edu/_common/files/pdf-climate/ucsb-full-report.pdf

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686. https://doi.org/10.21105/joss.01686
