---
title: "Parameter Estimation – Wild Fish Catch"
author: "Bri Baker"
date: "3/2/2021"
output:
  html_document:
    theme: cosmo
    code_folding: hide
    highlight: kate
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(nlstools)
library(broom)

```

## Summary

Wild fish catch has increased on a global scale. These data enumerate global wild fish catch in millions of tons per year from 1950 to 2012 (Earth Policy Institute, 2012). From the observed values, I created a model for the data utilizing nonlinear least squares. 


```{r data import}
fish <- read_csv(here("data/fish_catch.csv"),
                 skip = 2,
                 col_names = TRUE) %>% 
  clean_names() %>% 
  na.omit() %>% 
  rowid_to_column("years") %>% 
  mutate(years = years-1) %>% 
  mutate_if(is.character, as.numeric,) %>% 
  select(years, wild_catch)
```


## Data Exploration

Exploratory plot of wild caught fish.

```{r explore}
ggplot(fish, aes(years, wild_catch))+
  geom_point()+
  labs(x = "years", 
       y = "wild caught fish (million tons)")+
  scale_y_continuous(breaks = seq(0, 100, 10))+
  theme_minimal()

```

This exploratory plot indicates that the data roughly follow a logistic growth curve. This curve can be modeled as $P(t)=\frac{K}{1+Ae^{-kt}}$($K$ = carrying capacity, $A$ = $\frac{K-P_0}{P_0}$, $k$ = growth rate constant).

## Initial parameter estimation

```{r initial param}
K <-  90 # carrying capacity at ~90
A <- (K-10)/10 # Po ~10

exp_fish <- fish %>% 
  filter(years < 38) %>% # filter for exp
  mutate(ln_pop = log(wild_catch)) # log transform

lm_fish <- lm(ln_pop ~ years, data = exp_fish) # make linear model

r <- lm_fish$coefficients[[2]] # pull coefficient

```

Estimated starting parameters: $K$ = `r K`, $A$ = `r A`, $k$ = `r r`.


## Nonlinear least squares parameter estimation

```{r calc param}
fish_nls <- nls(wild_catch ~ K/(1 + A*exp(-r*years)), # run nls
                data = fish,
                start = list(K = K, A = A, r = r),
                trace = FALSE)

fish_model <- broom::tidy(fish_nls) #tidy to tibble

```

The calculated parameter variables are as follows: $K$ = `r round(fish_model$estimate[1], 2)` million tons, $A$ = `r round(fish_model$estimate[2], 2)` million tons, $r$ = `r round(fish_model$estimate[3], 2)` million tons/year. 


## Model visualization

```{r predict}

fish_predict <- predict(fish_nls) # make predictions for every year

fish_complete <- data.frame(fish, fish_predict) #bind to original df

# Plot them all together:
ggplot(data = fish_complete, aes(x = years, y = wild_catch)) +
  geom_point() +
  geom_line(aes(x = years, y = fish_predict)) +
  labs(title = "Global increase in wild fish catch from 1950 - 2012",
       subtitle = "Observed values and modeled growth curve",
       x = "Year",
       y = "Fish (million tons)",
       caption = "Bri Baker, 2021\n(Earth Policy Institute, 2012)") +
  scale_x_continuous(labels = seq(1950, 2010, 20)) +
  theme_minimal()

```



## Citations

Baty, Florent, Christian Ritz, Sandrine Charles, Martin Brutsche, Jean-Pierre Flandrois, Marie-Laure Delignette-Muller (2015). A Toolbox for Nonlinear Regression in R: The Package nlstools. Journal of Statistical Software, 66(5), 1-21. URL http://www.jstatsoft.org/v66/i05/.

Earth Policy Institute (2012). Global wild fish catch and aquaculture production, compiled by Earth Policy Institute with 1950-2010 from U.N. Food and Agriculture Organization (FAO), Global Capture Production and Global Aquaculture Production, electronic databases, at www.fao.org/fishery/topic/16140/en.

Firke, Sam (2020). janitor: Simple Tools for Examining and Cleaning Dirty Data. R package version 2.0.1. https://CRAN.R-project.org/package=janitor

Müller, Kirill (2020). here: A Simpler Way to Find Your Files. R package version 1.0.0. https://CRAN.R-project.org/package=here

Robinson, David, Alex Hayes and Simon Couch (2020). broom: Convert Statistical Objects into Tidy Tibbles. R package version 0.7.2. https://CRAN.R-project.org/package=broom

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686. https://doi.org/10.21105/joss.01686
